\thetitle

### 2021-03-03 writing/related.tex
organize sections/titles/labels in paper
related

\section{Related Work}

This paper builds on~\citet{lksfd-popl-2020}'s one-of-a-kind analysis of
blame for higher-order contracts. They do not spell out the notion of the
rational programmer but the basic ingredients of our technique derive
directly from theirs. Specifically, we adapt their empirical technique
to gradual typing which requires the development of custom mutators,
careful selection of debugging scenarios and generalization of the
technique to allow the comparison of different semantics.

The literature on gradual typing presents many semantics beyond the three
that we study, but only two additional blame strategies.
Pyret (\shorturl{https://www.}{pyret.org}) treats fixed-size data types same as Natural
and functions same as Transient. The Forgetful~\cite{cl-icfp-2017} and
the Amnesic~\cite{gfd-oopsla-2019} semantics are similar to Transient but
use wrappers instead of inlined checks.  Nom~\cite{mt-oopsla-2017} and
other \emph{concrete\/} semantics~\cite{wnlov-popl-2010, rsfbv-popl-2015,
rzv-ecoop-2015, rat-oopsla-2017} assume that every value comes with a type tag
and use tag checks to supervise the interactions between typed and untyped code.
Monotonic
references~\cite{svctg-esop-2015} and the semantics~\cite{tlt-popl-2019,
etg-icfp-19, tt-scp-20, tgt-popl-18, tt-sas-17} derived with the
Abstracting Gradual Typing
technique~\cite{gct-popl-2016} are (optimized) variants of Natural.
Among these representative semantics, only Amnesic and Nom present innovative
blame strategies.
Our study nevertheless excludes these strategies because Amnesic is a mere
theoretical construction and Nom imposes restrictions on untyped
code that would necessitate new data definitions across our benchmarks.

>>>>

%% -----------------------------------------------------------------------------

\citet{lksfd-popl-2020}'s human-out-of-the-loop empirical analysis of blame for
higher-order contracts in the inspiration for this paper.  While they do not
spell out the notion of the rational programmer, they present the basic
ingredients of the technique.

Technically, this work and~\citet{lksfd-popl-2020}'s differ in many ways. The
experiment presented here concerns three different modes of gradual typing,
theirs a single notion of contract checking.  It also contributes the idea of
creating three comparable rational programmers (with several modes). The
type-level mutators represent another technical contribution of the work
presented here; as explained in section~\ref{sec:mutate}, almost none
of~\citet{lksfd-popl-2020}'s mutators is useful at the level of type mistakes.

The literature on gradual typing presents many semantics beyond the three used
here and only two additional blame strategies.  Pyret
(\shorturl{https://www.}{pyret.org}) assigns fixed-size data types the Natural
semantics and functions a Transient semantics. The Forgetful~\cite{cl-icfp-2017}
and the Amnesic~\cite{gfd-oopsla-2019} semantics are similar to Transient but
use wrappers instead of in-lined checks.  Nom~\cite{mt-oopsla-2017} and other
\emph{concrete\/} semantics~\cite{wnlov-popl-2010, rsfbv-popl-2015,
rzv-ecoop-2015, rat-oopsla-2017} assume that every value comes with a type tag
and use tag checks to supervise the interactions between typed and untyped code.
Monotonic references~\cite{svctg-esop-2015} and the
semantics~\cite{tlt-popl-2019, etg-icfp-19, tt-scp-20, tgt-popl-18, tt-sas-17}
derived with the Abstracting Gradual Typing technique~\cite{gct-popl-2016} are
variants of Natural.  Among these representative semantics, only Amnesic and Nom
present innovative blame strategies.  The experiment excludes these strategies
because they seem to be merely theoretical constructions and/or impose severe
restrictions on programmers.
